---
title: "question2.1"
output: html_notebook
---


```{r}
data = get(load('../datasets/GoldenGate.RData'))
ts.gg<- ts(data[,'traffic'], start=1968, frequency=12)
```
```{r}
ggtsdisplay(ts.gg)
```

# a
For the time series selected generate time series which are corrected for trend, for seasonality and for
both trend and seasonality. Perform an Exploratory Data Analysis (including fingerprinting!) for the results!
```{r}
# remove trend
diff1.gg <- diff(ts.gg, differences = 1)
ggtsdisplay(diff1.gg)
```
```{r}
#remove seasonality

nsdiffs(ts.gg)
sd1.gg <- diff(ts.gg, lag=12, differences = 1)
ggtsdisplay(sd1.gg)
```

```{r}
# remove trend and seasonality

add.gg <- decompose(ts.gg, type='additive')
autoplot(add.gg)
```
```{r}
ggtsdisplay(add.gg$random)
```

# b
For the generated time series, corrected both for trend and seasonality select an adequate Exponential
Smoothing Model to predict future values.
Fit the model by estimating its parameters and give an interpretation for the parametervalues obtained.
Verify the model on the basis of the residuals, their (partial) autocorrelations and Goodness-of-Fit
measures such as RMSE, MAE and MAPE.
```{r}
trend <- diff(ts.gg, lag=12, differences = 1)
trend.season <- diff(trend, lag=12, differences = 1)
```


```{r}
#fc <- ses(ts.gg, h=15)
gg.ses1 <- HoltWinters(trend.season, alpha=0.2, beta=FALSE, gamma=FALSE)
autoplot(gg.ses1$fitted)
plot(gg.ses1)
```

```{r}
with(gg.ses1,accuracy(fitted,x))
```

```{r}
checkresiduals(gg.ses1)
```

# c
For the generated time series, corrected only for seasonality select an adequate Exponential Smoothing
Model to predict future values.
Fit the model by estimating its parameters and give an interpretation for the parametervalues obtained. Verify the model on the basis of the residuals, their (partial) autocorrelations and Goodness-of-Fit
measures such as RMSE, MAE and MAPE.

```{r}
season <- diff(ts.gg, lag=12, differences = 1)
gg.ses1 <- HoltWinters(season, alpha=0.2, beta=FALSE, gamma=FALSE)
plot(gg.ses1)
with(gg.ses1,accuracy(fitted,x))
checkresiduals(gg.ses1)
```

# d

For the generated time series, corrected only for trend select an adequate Exponential Smoothing Model
to predict future values.
Fit the model by estimating its parameters and give an interpretation for the parametervalues obtained.
Verify the model on the basis of the residuals, their (partial) autocorrelations and Goodness-of-Fit
measures such as RMSE, MAE and MAPE.

```{r}
trend <- diff(ts.gg, differences = 1)
gg.ses1 <- HoltWinters(trend, alpha=0.2, beta=FALSE, gamma=FALSE)
plot(gg.ses1)
with(gg.ses1,accuracy(fitted,x))
checkresiduals(gg.ses1)
```

# e

For the original time series, not corrected for trend and seasonality, select an adequate Exponential
Smoothing Model to predict future values.
Fit the model by estimating its parameters and give an interpretation for the parametervalues obtained.
Verify the model on the basis of the residuals, their (partial) autocorrelations and Goodness-of-Fit
measures such as RMSE, MAE and MAPE.

```{r}
gg.ses1 <- HoltWinters(ts.gg, alpha=0.2, beta=FALSE, gamma=FALSE)
plot(gg.ses1)
with(gg.ses1,accuracy(fitted,x))
checkresiduals(gg.ses1)
```

