---
title: "question6"
output: ''
---
# a.
For a number of time series realizations fingerprints in the time domain are given by means of their
autocorrelation functions. However, the order of presentations of time series and autocorrelation
functions is randomized. Identify the connection between the time series TS1, TS2, TS3 and the
autocorrelation functions AC1, AC2, AC3. Explicitly motivate your choice!

TS1:AC2 because it's the smoothest series, showing mostly just trend. It looks like the time series below, which has a similar AC plot.

```{r}
data = ts(get(load('../datasets/broadband.RData')))
ggtsdisplay(data[,'Market_1'], plot.type = 'spectrum')
```

TS3:AC1 because it shows pretty regular seasonality, which tends to appear in the AC plot as a clear up and down pattern that appear more drastic than a smoother timeseries like TS2. Therefore, TS2:AC3 because the auto correlation changes appear more smooth and is less significant. This lines up because TS2 seems to have less trend/cycle and seasonality than TS3, which has more significant ACs.

```{r}
data = ts(get(load('../datasets/catalog.RData')))

# similar to TS3:AC1
ggtsdisplay(data[,'men'], plot.type = 'spectrum')

# similar to TS2:AC3
ts.ma <- ma(data[,'men'], order = 5)
ggtsdisplay(ts.ma)
```

#b.
For a number of time series realizations the fingerprints in the frequency domain are given by means of
their smoothed periodograms. However, the order of presentations of time series and smoothed
periodograms is randomized. Identify the connection between the time series A, B, C and the smoothed
periodograms P1, P2, P3. Explicitly motivate your choice!

Time series C belongs to Periodigram P2 because this time series shows the least trend, which is captured in low frequencies of which P2 does not have much of.

P3 belongs to the time series with a cycle or seasonal pattern around frequency 0.5 and a smaller overall trend as spectrum = 20 when frequency = 0, unlike the higher spectrum value of 100 at frequency = 0 for P1. Because TSA seems to have a higher over all trend, I will pair it with P1, while that leaves TSC for P3.
