---
title: "question2"
output: html_notebook
---

In Exercise 1.1 you performed an Exploratory Data Analysis on some of the sample data sets. Now continue with a further analysis of these data sets selected by studying their fingerprints both in the time and the
frequency domain!

```{r}
data = get(load('../datasets/airline.RData'))
ts.airline <- ts(data$passenger,frequency=12, start=1949)
```

# a. 
Obtain ‘the fingerprint’ of the (original) time series in the time domain by calculating the sample
autocorrelation and partial autocorrelation function and comment on the results!

There is certainly strong AC and PAC with a lag of 12 months (weaker of 6 months)

```{r}
# Autocorrelation:
### Measure for the overall correlation between {xt-k} and {xt} Both through the direct and the indirect pathway
# Partial autocorrelation
### Measure for the overall correlation between {xt-k} and {xt} Both through the direct and the indirect pathway
ggtsdisplay(ts.airline)
ggAcf(ts.airline)
ggPacf(ts.airline)
```

# b. 
Apply finite and/or seasonal differencing to the original data to correct for trend and/or seasonality. Obtain
‘the fingerprint’ of the transformed time series in the time domain by calculating the sample
autocorrelation and partial autocorrelation function and comment on the results!

```{r}
# nearly gets rid of it
# differencing 1 for both is best
t1.diff <- diff(ts.airline, differences = 1)
t2.diff <- diff(ts.airline, differences = 2)
t1.s.diff <- diff(t1.diff, differences = 1, lag= 12)
t2.s.diff <- diff(t2.diff, differences = 1, lag= 12)
ggtsdisplay(t1.s.diff)
ggtsdisplay(t2.s.diff)
```



# c. 
Obtain ‘the fingerprint’ of both the original and the transformed time series in the frequency domain by
calculating the (smoothed) spectrum, using the “spectrum” program. Comment on the results and try to
identify frequency components that are present in the results!

```{r}
ggtsdisplay(ts.airline, plot.type = 'spectrum')
ggtsdisplay(t1.s.diff, plot.type = 'spectrum')
```

There are clearly homonics present in the data. There are many cycles reocurring every 6 months. Frequency components 2-6 are most present in the data.
Low frequences (trend) are not prominent.
